%!TEX root = ../main.tex
\section{Literature review}
\textcite{FF2015} introduce two additional factors to complement the \textcite{FamaFrench1993} three-factor asset pricing model. In what is referred to as the five-factor model, the traditional factors (market, value and size) are complemented by a profitability factor and an investment factor. Both factors represent zero-cost portfolios: the profitability factor, denoted RMW (robust-minus-weak), is long firms with high operating profitability and short firms with low operating profitability, and the investment factor, denoted CMA (conservative-minus-aggressive), is long firms with low investment rate and short firms with high investment rate. The five-factor model is found to be a siginificant improvement over the three-factor model, and the two new factors appear to have made the value factor, HML (high-minus-low), redundant. More specifically, the authors show that there is no significant intercept in regressions of HML on the remaining four factors, while each of the other factors have significant intercepts in similar regressions. The high average return of the HML factor appears to be fully explained by the remaining four factors. In an investment context, this can be interpreted as the HML factor adding no alpha to a portfolio holding the remaining four factors.

\textcite{Asness2015} challenge the notion that the value factor is subsumed by the addition of investment and profitability. In their study, they add a momentum factor, as well as an enhanced HML factor, and resurrect the alpha of the value factor. This leads us to believe that momentum could play an important role in recognizing the effect of HML. The momentum factor was originally studied by \textcite{JegadeeshTitman1993} and has since been shown to be present in many financial return series~\autocite{AsnessMoskovitzPedersen2013}.

It is debated whether factor strategies constitute rational risk premia or whether they are the consequences of market imperfections and irrational behavior. There are some appealing rational stories for the return premium of HML. \textcite{FamaFrench1993} show that the HML factor is related to systematic patterns of profitability and growth, and could proxy for a common risk source. This is supported by \textcite{LiewVassalou2000}, who show that the value factor can predict real GDP growth on data in several markets. \textcite{Zhang2005} uses a neoclassical model with rational expectations and competitive equilibrium to show that value firms have more tangible assets and are burdened by industry over-capacity in downturns, leading to higher down-market betas. \textcite{PetkovaZhang2005} find that the conditional betas of value stocks covary positively with the expected market risk premium.

Despite there being a number of rational theories, they all predict effects that are fairly small and cannot fully motivate the value premium. So far, the most pervasive explanations of the value premium are based on market imperfections and irrational behavior.\footnote{See i.a. \textcite{Ilmanen2011}}

\textcite{LakonishokShleiferVishny1994} argue that the value factor is driven by investors' overreaction to changes in earnings. They show that value firms have often experieced a decline in earnings over the last three years, lowering their book to market ratio. When earnings have gone down, investors as a group extrapolate the trend into the future and push prices away from fundamentals, giving rise to higher average returns for value firms and vice versa. Similar to \textcite{LakonishokShleiferVishny1994}, \textcite{BarberisHuang2001} draw on the fact that value firms have experienced decreasing earnings, but suggest that the premium is driven by investors' loss aversion bias. Current value firms have had falling earnings, leading to lower share prices and negative returns. As investors are deterred by the past performance of negative returns in itself, they are less willing to hold value stocks. \textcite{LakonishokShleiferVishny1992} instead find an explanation in the structure of the money manager industry. They suggest that there money managers have career based incentives to avoid value stocks as such stocks are more likely to go bankrupt, and make the short-term performance look bad, than growth stocks, which are more widely held in the reference index.

Variations of the long-short strategy of the value factor has become a staple strategy of both quantitative and qualitative hedge funds, often under the "equity market neutral" or "fundamental quantitative" labels. Factor equity strategies have also become increasingly accessible for retail investors, especially with the advent of smart beta ETFs. \footnote{See i.a. \textcite{Pedersen2015}, \textcite{AQREMN} and \textcite{McKEMN}.} 

The use of leverage in hedge funds can exacerbate the flow patterns in factor strategies, as highlighted by the quant crash in July-August 2007. \textcite{KhandaniLo2011} and \textcite{KhandaniLo2007} revisit the sudden and large losses of factor strategies (including value and size) during this period, and provide evidence for the "Unwind hypothesis": The crash started with rapid sell-offs of large blocks of factor strategy portfolios, for which there was not enough liquidity to maintain prices. The price drops, in turn, led to further liquidations due to 1) margin calls in other leveraged and long-short funds and 2) risk management policies, even in traditional long-only funds. This liquidity and margin spiral is very similar to that proposed by \textcite{Brunnermeier2009} and \textcite{BrunnermeierPedersen2009}.

Other papers have also highlighted the risk of "crowded trades" where leverage is applied. \textcite{Stein2009} shows that markets can become less pricing efficient and have increased chances of large fire-sale crashes when rational investors set their leverage level unknowing of how many others are engaging in a similar trade. \textcite{LouPolk2013} introduce measures of arbitrage activity and show that momentum strategies become destabilizing and prone to crash in times of high activity. For the value strategy, high arbitrage activity is instead shown to forecast positive value returns. 

The profitability and investment factors have only recently made their way into academic literature. \textcite{NovyMarx2013} study the return differences between firms with high and low gross profitability and shows significant abnormal returns to a profitability factor. Profitability is also shown to have approximately the same power in predicting the cross-section of stock returns as does value (book-to-market). Furthermore, the profitability strategy is negatively correlated with value, and can therefore improve the investing performance of a value strategy. \textcite{CooperGulenSchill2008} investigate investment, measured as the percentage change in total assets, and show that the related zero-cost-portfolio provides significant abnormal returns and that it has additional predictive ability in the cross-section of stock returns, taking both value and size into account.

While all other factor pairs are nearly uncorrelated or negatively correlated, the value (HML) and investment (CMA) factors are highly positively correlated. \textcite{Zhang2005} predicts this positive relation in a model setting, and \textcite{AndersonGarciaFeijoo2006} confirm it on empirical data. More specifically, the empirical study shows that past investment has a significant positive relation with the book-to-market ratio. In other words, value firms with high book-to-market might be value firms precisely because they have invested little, and vice versa. \textcite{FF2015} consider it a fact that value firms invest less than growth firms.

There are only a handful of papers that study factor strategies using copula methods. A working paper by \textcite{CholleteNing2012} examine dynamic correlations between a four-factor model and aggregate US consumption, and find evidence for tail dependence across the five risk factors. \textcite{ChristoffersenLanglois2013} study the four-factor model alone on US data 1963-2010, and show significant and asymmetric tail dependence that cannot be captured by standard linear correlation measures. A skewed Student-\textit{t} copula-GARCH model is found to be able to generate the model fairly well, and the authors proceed with 20 years of out-of-sample analysis on investing based on conditional expectations from the copula model, leading to significant improvements for investors with a CRRA utility function. To the best of our knowledge, no paper has previously studied the role of the investment and profitability factors using copula models.