%!TEX root = ../main.tex

\section{Modeling of Factor Returns} % (fold)
\label{sec:modeling_of_factor_returns}

% Why a model?
This section presents our model for the joint behavior of returns. A multivariate model of returns allows us to make conditional forecasts of the distribution of returns one week ahead, which take into account the dependence between factors. The model is used in the mean-variance analysis, to provide dynamic inputs that can give us optimal weights over time. The model is also used in the analysis of diversification benefits, when we shift the focus to the tail risk of factor portfolios. First, we describe why we choose the copula model among different multivariate models. Second, we define the model and interpret the parameterization. Third, we select and estimate univariate models that are building blocks in the copula. Fourth, we analyze residuals from univariate models, to determine what type of multivariate dependence the copula should capture. Fifth, we estimate the copula and choose the best specification. Last, we conduct a robustness check of whether the copula captures the dependence structure.

\input{tex/model/choosing} % motivating the choice of copula, and among copulas

\input{tex/model/copula} % the boom

\input{tex/model/univariate} % arma garch

\input{tex/model/dependence} % threshold and rolling

\input{tex/model/results} % estimation results and explaining parameterization choice

\input{tex/model/robustness} % constant threshold corrs simulated, maybe also add robustness check of whether the model captures roll-corrs 

% section modeling_of_factor_returns (end)
