%!TEX root = ../main.tex

\section{Optimizing factor allocations} % (fold)
\label{sec:optimizing_factor_allocations}

We now turn to the issue of optimizing factor allocations using our estimated model. Large money managers in the factor space generally advice against factor timing and use static strategies that are \emph{based} on equal-weighting -- a simple heuristic that has proven hard to beat out-of-sample. The managers do, however, blend the equal-weights approach with optimization routines, such as mean-variance and minimum-volatility, to arrive at static policy weights.\footnote{See i.a. \textcite{AQRSiren}, \textcite{BlackRock}, \textcite{MSCI} and \textcite{Robeco}.}

Our research question concerns the importance of HML given the discovery of RMW and CMA. Based on abnormal return regressions and the work on mean-variance portfolios in \textcite{HubermanKandel1987}, \textcite{FF2015} conjecture that the HML factor will not improve the mean-variance tangency portfolio when added to a portfolio of Mkt.RF, SMB, RMW and CMA. Differently put, this is equivalent to HML having a zero weight in mean-variance optimization. We evaluate this both based on static weights (based on sample estimates of expected returns and covariances) and dynamic weights (based on copula model estimates).

While dynamic weights may seem counter-intuitive at first, as investors use static allocations, we argue that the dynamic analysis is essential. Investors carry out real-time investing and can only use information up until the current date. We, on the other hand, evaluate the \emph{ex-post} merits of different asset universes (e.g. including and excluding HML), and use all available data to determine dynamic weights.

Moving beyond the mean-variance framework, because factor returns are not normal, the covariance matrix used in mean-variance optimization is not a full description of the dependency between factors.

This leads us to consider optimization using two techniques: Optimal Mean-Variance (MV) and optimal Conditional Diversification Benefit (CDB), which is a measure introduced by~\textcite{ChristoffersenErrunzaJacobLanglois2012}. The first optimization is a direct test of the conjecture in~\textcite{FF2015} that HML does not expand the mean-variance frontier. The second optimization is done to test the relative diversification benefit of different factors -- HML, CMA, RMW -- under non-normality.

CDB is based on the portfolio expected shortfall (ES), i.e. the expected loss in case the return realizes below its Value-at-Risk (VaR), and therefore concerns the properties of the lower tail of the portfolio distribution. This allows it to consider non-normalities in the return distribution.

The remainder of this section presents the construction of the CDB measure and presents the common optimization problem. Results of the optimizations are presented in the upcoming sections.

\subsection{Conditional Diversification Benefit} % (fold)
\label{sub:conditional_diversification_benefit}

This description of CDB follows~\textcite{ChristoffersenErrunzaJacobLanglois2012}. Define ES as the expected loss in some bottom percentile $q$:
\begin{align}
    \text{ES}_{i,t}^q(r_{i,t}) = -\mathbb{E}[r_{i,t} | r_{i,t} \leq F_{i,t}^{-1}(q)]
\end{align}
where $F_{i,t}^{-1}(q)$ is the inverse CDF of simple returns $r_{i,t}$ at $q$ (equivalent to the $q\%$ Value-at-Risk). 

The expected shortfall represents the expected loss when returns realize below the Value-at-Risk of the portfolio. Depending on the distribution at hand, the expected shortfall can be closer or further to the Value-at-Risk. Intuitively, if assets offer little diversification, then no combination of assets will reduce total portfolio risk; and ES will be higher. 

For a portfolio of assets with weights $w_t$, the portfolio ES $\text{ES}_t^q(w_t)$ has an upper bound equal to the weighted average of each asset's ES, corresponding to the case of no diversification~\autocite{Artzner1999}:
\begin{align}
  \overline{\text{ES}}_t^q(w_t) = \sum_{i=1}^N w_{i,t} \text{ES}_{i,t}^q(r_{i,t})
\end{align}
A lower bound on portfolio ES is given by the portfolio's Value-at-Risk ($-F_{t}^{-1}(w_t, q)$):
\begin{align}
  \underline{\text{ES}}_t^q(w_t) = -F_{t}^{-1}(w_t, q)
\end{align}
CDB is defined as the portfolio's ES scaled by its lower and upper bounds:
\begin{align}
  \text{CDB}_t^q(w_t) = \frac{\overline{\text{ES}}_t^q(w_t) - \text{ES}_t^q(w_t)}{\overline{\text{ES}}_t^q(w_t) - \underline{\text{ES}}_t^q(w_t)}
\end{align}
CDB is a number between zero and one (we report CDB scaled to 0--100) and thus measures how close to its Value-at-Risk the expected shortfall of a portfolio gets (we report CDB scaled by 100). Note that the level of expected return does not enter into the measure -- CDB only measures how ``useful'' a group of assets are for eliminating risk.

% subsection conditional_diversification_benefit (end)

\subsection{Optimization Problem} % (fold)
\label{sub:optimization_problem}

Each week, we choose portfolio weights to maximize the Sharpe Ratio (SR) in the mean-variance case, and CDB in the CDB case. We impose two restrictions on our optimization problem. First, all factor weights must be positive. This reduces the problem with extreme weights, seen in the unconstrained optimization problem, and reflects a view that an investor will not bet against factors that have generated a history of positive premia. Second, factor weights must sum to unity, i.e. fully invested across factors. In light of~\textcite{Asness2015}, we consider portfolios with and without momentum (five- and six-factor portfolios, respectively) and how portfolio weights change when replacing HML with CMA and vice versa.

Together, these restrictions make the maximized Sharpe Ratio reflect tangency weights subject to a constraint of no negative weights. Due to the restrictions, the standard analytical solution to the mean-variance problem is not equal to the tangency portfolio. Similarly, no analytical solution exists for optimizing CDB. Hence, we performa a numerical optimization where we choose weights $w_t$ to maximize the objective function $\Omega_t(w_t)$ subject to the above restrictions:
\begin{align}
  \arg\!\max_{w_t} \Omega_t(w_t)
    &&\text{\quad s.t.\quad} & \sum_{i=1}^N w_{i,t} = 1 \\
    &&                       & w_{i,t} \ge 0 \notag
\end{align}
For the MV case, the objective function is the one-week Sharpe Ratio (SR):
\begin{align}
  \Omega_t(w_t) = \frac{w_t^\top \mathbb{E}_t[r_{t+1}]}{\sqrt{w_t^\top \mathbb{E}_t[\Sigma_{t+1}] w_t}}
\end{align}
$\mathbb{E}_t[r_{t+1}]$ is the conditional one-step-ahead expected factor return and $\mathbb{E}_t[\Sigma_{t+1}]$ is the conditional one-step-ahead variance-covariance matrix. 

For the CDB case, the objective function is the one-week CDB:
\begin{align}
  \Omega_t(w_t) = \text{CDB}_t^q(w_t)
\end{align}
Each week, we perform MV and CDB optimization using simulated distributions from our fully estimated dynamic copula model. Each week, we simulate $10\,000$ returns from the estimated copula model and use those outcomes as the conditional distribution of factor returns. We also perform MV optimization based on sample means and covariances, in which case $\mathbb{E}_t[r_{t+1}]$ and $\mathbb{E}_t[\Sigma_{t+1}]$ are constant and equal to the full-sample estimates.

% subsection optimization_problem (end)

% section optimizing_factor_allocations (end)
